---
title: "Report"
author: "Angela Zhai"
date: "10/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(knitr, tidyverse)
```

## I. Introduction

The vehicle level crash data comes from Massachusetts government website (https://www.mass.gov/). When car crashes are reported to the Registry of Motor Vehicles, the MassDOT Highway Division uses the crash information for data-driven safety analyses and traffic engineering studies. We are interested to figure out if any factors related to the crashes.



## II. Data

```{r echo=FALSE}
## Import Crash Data
Crash <- read_csv("2017_Vehicle_Level_Crash_Details_.csv",
                  col_types = cols(EXIT_NUMB = col_character(), EXIT_RTE = col_character(), MM_RTE = col_character()))
```


### Overview

```{r echo=FALSE}
kable(Crash[1:5, ])
```

### Exploratory Data Analysis

```{r echo=FALSE}
# colnames(Crash)
## There are 138 columns in the crash data
```

```{r echo=FALSE}
## Only keep interested fileds, and order by IDs
Crash_clean <- Crash %>%
  select(OBJECTID, CRASH_NUMB, RMV_DOC_IDS, CRASH_RPT_IDS, YEAR, CRASH_DATE, CRASH_HOUR, CITY_TOWN_NAME,
         CRASH_SEVERITY_DESCR, NUMB_VEHC, AMBNT_LIGHT_DESCR, WEATH_COND_DESCR, ROAD_SURF_COND_DESCR,
         RDWY_JNCT_TYPE_DESCR, TRAF_CNTRL_DEVC_TYPE_DESCR, SPEED_LIMIT, URBAN_TYPE, URBAN_LOC_TYPE, DRIVER_AGE,
         MOST_HRMF_EVENT) %>%
  arrange(CRASH_NUMB, RMV_DOC_IDS)
## Seems like if the crash happened by multiple vehicles, we have separated record for each of them.
```

```{r echo=FALSE}
## Only keep single record for single crash
Crash_event <- Crash_clean %>%
  select(-OBJECTID, -DRIVER_AGE, -MOST_HRMF_EVENT) %>%
  distinct()
```


Question: Was it light or dark when crash happened?

```{r echo=FALSE}
Crash_event %>%
  group_by(AMBNT_LIGHT_DESCR) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

Question: What's the most severe crash (most vehicles got involved) under each weather condition? Show top 10 here.

```{r echo=FALSE}
Crash_event %>%
  group_by(WEATH_COND_DESCR) %>%
  summarise(max_num_vehicle = max(NUMB_VEHC)) %>%
  arrange(desc(max_num_vehicle)) %>%
  head(10)
```


Visualization: Total Number of crashes happened by hour

```{r echo=FALSE}
Crash_hourly <- Crash_event %>%
  group_by(CRASH_HOUR) %>%
  summarise(count = n()) %>%
  filter(CRASH_HOUR != '')
```

```{r echo=FALSE}
Crash_hourly$CRASH_HOUR <- factor(Crash_hourly$CRASH_HOUR,
                                  levels = c("12:00AM to 12:59AM", "01:00AM to 01:59AM", "02:00AM to 02:59AM",
                                             "03:00AM to 03:59AM", "04:00AM to 04:59AM", "05:00AM to 05:59AM",
                                             "06:00AM to 06:59AM", "07:00AM to 07:59AM", "08:00AM to 08:59AM",
                                             "09:00AM to 09:59AM", "10:00AM to 10:59AM", "11:00AM to 11:59AM",
                                             "12:00PM to 12:59PM", "01:00PM to 01:59PM", "02:00PM to 02:59PM",
                                             "03:00PM to 03:59PM", "04:00PM to 04:59PM", "05:00PM to 05:59PM",
                                             "06:00PM to 06:59PM", "07:00PM to 07:59PM", "08:00PM to 08:59PM",
                                             "09:00PM to 09:59PM", "10:00PM to 10:59PM", "11:00PM to 11:59PM"))
```


```{r echo=FALSE}
ggplot(data = Crash_hourly) +
  geom_point(mapping = aes(CRASH_HOUR, count), size=5, col='orange') +
  geom_line(mapping = aes(CRASH_HOUR, count), group=1, size=3) +
  theme(axis.text.x = element_text(angle = 30))
```


